---
title: "Script to check accuracy and % of missing EUNIS for EVA and EVA Resurvey"
output:
  pdf_document: default
  html_notebook: default
---

```{r, include=FALSE}
# Set global chunk options
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
knitr::opts_chunk$set(fig.width=5.5, fig.height=3.5)
```

```{r, include=FALSE}
# Load libraries
library(here)
library(readr)
library(tidyr)
library(dplyr)
library(lubridate)
library(ggplot2)
library(stringr)
library(purrr)
library(readxl)
library(gridExtra)
library(cowplot)
```

```{r, include=FALSE}
# Define functions
printall <- function(tibble) {
  print(tibble, width = Inf)
  }
```

```{r, include=FALSE}
# Read the data
db_resurv_updated_clean<-read_tsv(
  here("data", "clean", "db_resurv_updated_clean.csv"),
  col_types = cols(
    # Dynamically specify EUNIS columns as character
    .default = col_guess(),  # Default guessing for other columns
    EUNISa = col_character(),
    EUNISb = col_character(),
    EUNISc = col_character(),
    EUNISd = col_character(),
    EUNISa_1 = col_character(),
    EUNISa_2 = col_character(),
    EUNISa_3 = col_character(),
    EUNISa_4 = col_character(),
    EUNISb_1 = col_character(),
    EUNISb_2 = col_character(),
    EUNISb_3 = col_character(),
    EUNISb_4 = col_character(),
    EUNISc_1 = col_character(),
    EUNISc_2 = col_character(),
    EUNISc_3 = col_character(),
    EUNISc_4 = col_character(),
    EUNISd_1 = col_character(),
    EUNISd_2 = col_character(),
    EUNISd_3 = col_character(),
    EUNISd_4 = col_character(),
    EUNISa_1_descr = col_character(),
    EUNISc_1_descr = col_character(),
    EUNISd_1_descr = col_character(),
    EUNIS_assignation = col_character(),
    EUNISa_2_descr = col_character(),
    EUNISa_3_descr = col_character(),
    EUNISa_4_descr = col_character(),
    EUNISb_2_descr = col_character(),
    EUNISb_3_descr = col_character(),
    EUNISb_4_descr = col_character(),
    EUNISc_2_descr = col_character(),
    EUNISc_3_descr = col_character(),
    EUNISc_4_descr = col_character(),
    EUNISd_2_descr = col_character(),
    EUNISd_3_descr = col_character(),
    EUNISd_4_descr = col_character()
    )
  )
db_EVA_updated_clean<-read_tsv(
  here("data", "clean", "db_EVA_updated_clean.csv"),
  col_types = cols(
    # Dynamically specify EUNIS columns as character
    .default = col_guess(),  # Default guessing for other columns
    EUNISa = col_character(),
    EUNISb = col_character(),
    EUNISc = col_character(),
    EUNISd = col_character(),
    EUNISa_1 = col_character(),
    EUNISa_2 = col_character(),
    EUNISa_3 = col_character(),
    EUNISa_4 = col_character(),
    EUNISb_1 = col_character(),
    EUNISb_2 = col_character(),
    EUNISb_3 = col_character(),
    EUNISb_4 = col_character(),
    EUNISc_1 = col_character(),
    EUNISc_2 = col_character(),
    EUNISc_3 = col_character(),
    EUNISc_4 = col_character(),
    EUNISd_1 = col_character(),
    EUNISd_2 = col_character(),
    EUNISd_3 = col_character(),
    EUNISd_4 = col_character(),
    EUNISa_1_descr = col_character(),
    EUNIS_assignation = col_character(),
    EUNISa_2_descr = col_character(),
    EUNISa_3_descr = col_character(),
    EUNISa_4_descr = col_character(),
    EUNISb_2_descr = col_character(),
    EUNISb_3_descr = col_character(),
    EUNISb_4_descr = col_character(),
    EUNISc_2_descr = col_character(),
    EUNISc_3_descr = col_character(),
    EUNISc_4_descr = col_character(),
    EUNISd_2_descr = col_character(),
    EUNISd_3_descr = col_character(),
    EUNISd_4_descr = col_character(),
    Lon_prec = col_double(),
    Lat_prec = col_double(),
    EUNISc_1_descr = col_character(),
    EUNISd_1_descr = col_character(),
    Private = col_character()
  )
)
# No parsing issues!
```

```{r, include=FALSE}
# Select key columns and add "database" column
db_resurv_updated_clean <- db_resurv_updated_clean %>%
  select(PlotObservationID, Country, Altitude, `Aspect (°)`, `Slope (°)`,
         starts_with("EUNIS"), `ReSurvey project`, `ReSurvey site`,
         `ReSurvey plot`, `ReSurvey observation`, `Manipulate (y/n)`,
         `Type of manipulation`, `Location method`, RS_CODE, RS_PROJTYP, 
         RS_DUPL, `Location uncertainty (m)`, Dataset, precision_new,
         Lon_updated, Lat_updated, coordinates_equal, coordinates_consistent,
         n_EUNIS, date, year, EUNIS_assignation) %>%
  mutate(database = "EVA_RS")
```

```{r, include=FALSE}
db_EVA_updated_clean <- db_EVA_updated_clean %>%
  select(PlotObservationID, Country, Altitude, `Aspect (°)`, `Slope (°)`,
         starts_with("EUNIS"), `ReSurvey project`, `ReSurvey site`,
         `ReSurvey plot`, `ReSurvey observation`, `Manipulate (y/n)`,
         `Type of manipulation`, `Location method`, RS_CODE, RS_PROJTYP, 
         RS_DUPL, `Location uncertainty (m)`, Dataset, precision_new,
         Lon_updated, Lat_updated, # coordinates_equal, coordinates_consistent,
         n_EUNIS, date, year, EUNIS_assignation) %>%
  mutate(database = "EVA") %>%
  mutate(coordinates_equal = NA, coordinates_consistent = NA)
```

```{r, include=FALSE}
# Combine the two databases
db_resurv_and_EVA <- bind_rows(db_resurv_updated_clean, db_EVA_updated_clean)
```

# EUNIS level 1 availability

```{r}
ggplot(db_resurv_and_EVA %>% 
         # Create column to show availability of EUNIS level 1 (Y/N)
         mutate(EUNIS_1_available = ifelse(is.na(EUNISa_1), "N", "Y")), 
       aes(EUNIS_1_available, fill = database)) +
  geom_bar() +
  labs(y = "Number of observations") +
  ggtitle("Availability of EUNIS level 1 in EVA + EVA ReSurvey")
```

# EUNIS assignation

```{r}
ggplot(db_resurv_and_EVA, aes(EUNIS_assignation, fill = database)) +
  geom_bar() +
  labs(y = "Number of observations") +
  ggtitle("EUNIS assignation")
```

# Coordinates available

```{r}
ggplot(db_resurv_and_EVA %>% 
         # Create column to show availability of coordinates
         mutate(coords_available = 
                  ifelse(is.na(Lon_updated) & is.na(Lat_updated), "N", "Y")),
       aes(coords_available, fill = database)) +
  geom_bar() +
  labs(y = "Number of observations") +
  ggtitle("Availability of coordinates")
```

# Location uncertainty 

## Availabilty

```{r}
ggplot(db_resurv_and_EVA %>% 
         # Create column to show availability of location uncertainty
         mutate(loc_uncertainty_available = 
                  ifelse(is.na(`Location uncertainty (m)`), "N", "Y")),
       aes(loc_uncertainty_available, fill = database)) +
  geom_bar() +
  labs(y = "Number of observations") +
  ggtitle("Availability of location uncertainty")
```

## Distribution when available

```{r}
ggplot(db_resurv_and_EVA,
       aes(`Location uncertainty (m)`, fill = database)) +
  geom_histogram(aes(y = (..count..) / sum(..count..) * 100),
                 color = "black", binwidth = 10000) +
  xlab("Location uncertainty (m)") + ylab("% of observations") +
  ggtitle("Histogram with % of observations")
```

```{r}
ggplot(db_resurv_and_EVA %>% 
         # Create column to show two categories for location uncertainty
         mutate(
           Location_uncertainty_range = factor(
             case_when(
               `Location uncertainty (m)` <= 30 ~ "0 – 30 m",
               `Location uncertainty (m)` <= 200 ~ "30 m – 200 m",
               `Location uncertainty (m)` <= 1000 ~ "200 m - 1 km",
               `Location uncertainty (m)` <= 5000 ~ "1 km - 5 km",
               `Location uncertainty (m)` <= 10000 ~ "5 km - 10 km",
               is.na(`Location uncertainty (m)`) ~ "NA",
               TRUE ~ "10 km +"
               ), 
             levels = c("0 – 30 m", "30 m – 200 m", "200 m - 1 km", 
                        "1 km - 5 km", "5 km - 10 km", "10 km +", "NA"))
           ),
       aes(Location_uncertainty_range, fill = database)) +
  geom_bar(aes(y = (..count..) / sum(..count..) * 100)) +
  labs(y = "% of observations") +
  ggtitle("Range of location uncertainty (EVA + EVA ReSurvey)")
```

```{r}
ggplot(db_resurv_and_EVA %>%
         filter(`Location uncertainty (m)` < 10000),
       aes(`Location uncertainty (m)`, fill = database)) +
  geom_histogram(aes(y = (..count..) / sum(..count..) * 100), color = "black") +
  xlab("Location uncertainty (m)") + ylab("% of observations") +
  ggtitle("Histogram with % of observations (Loc. uncertainty < 10 km)")
```

```{r}
ggplot(db_resurv_and_EVA %>%
         filter(`Location uncertainty (m)` < 500),
       aes(`Location uncertainty (m)`, fill = database)) +
  geom_histogram(aes(y = (..count..) / sum(..count..) * 100), color = "black") +
  xlab("Location uncertainty (m)") + ylab("% of observations") +
  ggtitle("Histogram with % of observations (Loc. uncertainty < 0.5 km)")
```

## Location method

```{r fig.height=10, fig.width=13}
ggplot(db_resurv_and_EVA %>% 
         # Create column to show two categories for location uncertainty
         mutate(
           Location_uncertainty_range = factor(
             case_when(
               `Location uncertainty (m)` <= 30 ~ "0 – 30 m",
               `Location uncertainty (m)` <= 200 ~ "30 m – 200 m",
               `Location uncertainty (m)` <= 1000 ~ "200 m - 1 km",
               `Location uncertainty (m)` <= 5000 ~ "1 km - 5 km",
               `Location uncertainty (m)` <= 10000 ~ "5 km - 10 km",
               is.na(`Location uncertainty (m)`) ~ "NA",
               TRUE ~ "10 km +"
               ), 
             levels = c("0 – 30 m", "30 m – 200 m", "200 m - 1 km", 
                        "1 km - 5 km", "5 km - 10 km", "10 km +", "NA")),
           location_method = ifelse(
             `Location method` %in% c("04", "01", "06", "02", "08", "07", "05",
                                      "2", "03", "4"), 
             "Codes", `Location method`)
           ),
       aes(Location_uncertainty_range, fill = location_method)) +
  geom_bar(aes(y = (..count..) / sum(..count..) * 100), 
           position = position_dodge(width = 0.9)) +
  facet_wrap(~ database, scales = "free") +
  labs(y = "% of observations") +
  theme(legend.position = "top") +
  ggtitle("Range of location uncertainty (EVA + EVA ReSurvey)")
```

# Check for plots with multiple records for the same year

Careful! There are plots with muliple records for the same year (and sometimes same date).

```{r}
db_resurv_mult_obs_per_year <- db_resurv_and_EVA %>%
  filter(database == "EVA_RS") %>%
  filter(!is.na(`ReSurvey plot`)) %>%
  group_by(RS_CODE, `ReSurvey site`, `ReSurvey plot`, `year`) %>%
  summarise(count = n()) %>%
  filter(count > 1) %>%
  ungroup() %>%
  left_join(db_resurv_and_EVA, 
             by = c("RS_CODE", "ReSurvey site", "ReSurvey plot", "year")) 
```

```{r}
nrow_db_resurv_mult_obs_per_year <- nrow(
  db_resurv_mult_obs_per_year %>%
    distinct(RS_CODE, `ReSurvey site`, `ReSurvey plot`, year)
  )
```

For `r nrow_db_resurv_mult_obs_per_year` unique combinations of RS_CODE, ReSurvey site, ReSurvey plot and year, there are multiple records, sometimes with different dates within the year (OK - we choose one), **but also with the same exact date**.

Example below:

```{r}
data.frame(
  db_resurv_mult_obs_per_year %>%
    filter(`ReSurvey plot` == "N_Cep" & year ==1952) %>%
    select(PlotObservationID, RS_CODE, `ReSurvey site`, `ReSurvey plot`, year,
           date, count, Country, EUNISa, `ReSurvey project`,
           `ReSurvey observation`, Lon_updated, Lat_updated)
  )
```

I have sent a list to Ilona to see what to do with this issue.

```{r, include=FALSE}
# Save in file (fewer columns for internal use)
write_tsv(db_resurv_mult_obs_per_year %>%
            select(RS_CODE, `ReSurvey site`, `ReSurvey plot`,
                   `ReSurvey observation`, date, year, count,
                   EUNIS_assignation, EUNISa, EUNISb, EUNISc, EUNISd,
                   Lon_updated, Lat_updated, `Location uncertainty (m)`),
          here("data", "edited","db_resurv_mult_obs_per_year.csv"))
```

```{r, include=FALSE}
# Save in file (all columns to send to Ilona)
write_tsv(db_resurv_mult_obs_per_year,
          here("data", "edited","db_resurv_mult_obs_per_year_Ilona.csv"))
```

# Keep GOOD observations WP3

Defining "good" observations as those that **have an EUNIS assigned** (any level) and where **location uncertainty is available and < 0.5 km** (decide if we use 0.2 or other value?).

```{r, include=FALSE}
db_resurv_and_EVA_good <- db_resurv_and_EVA %>%
  filter(EUNIS_assignation != "Not possible" &
           `Location uncertainty (m)` < 500) 
```

```{r}
prop_good <- nrow(db_resurv_and_EVA_good) / nrow(db_resurv_and_EVA)
```

```{r}
prop_good_EVA_RS <- nrow(
  db_resurv_and_EVA_good %>% filter(database == "EVA_RS")
  ) /
  nrow(
    db_resurv_and_EVA %>% filter(database == "EVA_RS"))
```

From **EVA + EVA ReSurvey**, the **proportion of "good" observations** according to these criteria is **`r round(prop_good, 3)`**.

```{r}
prop_EUNIS <- nrow(
  db_resurv_and_EVA %>% filter(EUNIS_assignation != "Not possible")
  ) /
  nrow(db_resurv_and_EVA)
```

From **EVA ReSurvey only**, the **proportion of "good" observations** according to these criteria is **`r round(prop_good_EVA_RS, 3)`**.

```{r}
prop_EUNIS <- nrow(
  db_resurv_and_EVA %>% filter(EUNIS_assignation != "Not possible")
  ) /
  nrow(db_resurv_and_EVA)
```

```{r}
prop_EUNIS_EVA_RS <- nrow(
  db_resurv_and_EVA %>%
    filter(database == "EVA_RS" & EUNIS_assignation != "Not possible")
  ) /
  nrow(db_resurv_and_EVA %>% filter(database == "EVA_RS"))
```

From **EVA + EVA ReSurvey**, the **proportion of observations with EUNIS assigned, irrespective of location uncertainty** is **`r round(prop_EUNIS, 3)`**.

From **EVA ReSurvey only**, the **proportion of observations with EUNIS assigned, irrespective of location uncertainty** is **`r round(prop_EUNIS_EVA_RS, 3)`**.

Maybe include those without location uncertainty but where location method is GPS, etc.?

# Check for plots with multiple records for the same year in "good" observations

```{r}
db_resurv_mult_obs_per_year_good <- db_resurv_and_EVA_good %>%
  filter(database == "EVA_RS") %>%
  filter(!is.na(`ReSurvey plot`)) %>%
  group_by(RS_CODE, `ReSurvey site`, `ReSurvey plot`, `year`) %>%
  summarise(count = n()) %>%
  filter(count > 1) %>%
  ungroup() %>%
  # Add the column ReSurvey observation and EUNIS columns
  # to see how to proceed
  left_join(db_resurv_and_EVA, 
             by = c("RS_CODE", "ReSurvey site", "ReSurvey plot", "year")) %>%
  select(RS_CODE, `ReSurvey site`, `ReSurvey plot`,
         `ReSurvey observation`, date, year, count,
         EUNIS_assignation, EUNISa, EUNISb, EUNISc, EUNISd,
         Lon_updated, Lat_updated, `Location uncertainty (m)`)
```

# Keep GOOD observations that do not have multiple records for the same year

```{r}
db_resurv_and_EVA_good_no_multiples <- db_resurv_and_EVA_good %>%
  # Get the rows in db_resurv_and_EVA_good that do not have matching rows in
  # db_resurv_mult_obs_per_year_good based on the specified columns
  # (i.e. the rows that do not have multiple records for the same year)
  anti_join(
    db_resurv_mult_obs_per_year_good %>%
      select(RS_CODE, `ReSurvey site`, `ReSurvey plot`) %>%
      distinct(),
    by = c("RS_CODE", "ReSurvey site", "ReSurvey plot")
  )
```

```{r}
prop_good_no_multiples <- nrow(db_resurv_and_EVA_good_no_multiples) /
  nrow(db_resurv_and_EVA)
```

```{r}
prop_good_no_multiples_EVA_RS <- nrow(
  db_resurv_and_EVA_good_no_multiples %>% filter(database == "EVA_RS")
  ) /
  nrow(db_resurv_and_EVA %>% filter(database == "EVA_RS"))
```

From **EVA + EVA ReSurvey**, the **proportion of "good" observations** according to the above criteria **that do not have multiple records for the same year** is **`r round(prop_good_no_multiples, 3)`**.

From **EVA ReSurvey only**, the **proportion of "good" observations** according to the above criteria **that do not have multiple records for the same year** is **`r round(prop_good_no_multiples_EVA_RS, 3)`**.

# Location uncertainty (last resurveys for EVA_RS)

Extract last resurvey for each each ReSurvey plot (defined by combination of RS_CODE, ReSurvey site and ReSurvey plot). So far, discard 111 resurvey observations corresponding to ReSurvey project “Swiss arable flora since the 1920s” (RS_CODE = CH_0002) where ReSurvey plot is NA. Discard also plots that have multiple records for the same year (if keeping those, we have to select one record - randomly?).

```{r}
db_resurv_last <- db_resurv_and_EVA %>%
  # Get the rows in db_resurv_and_EVA that do not have matching rows in
  # db_resurv_mult_obs_per_year based on the specified columns
  # (i.e. the rows that do not have multiple records for the same year)
  anti_join(
    db_resurv_mult_obs_per_year %>%
      select(RS_CODE, `ReSurvey site`, `ReSurvey plot`) %>%
      distinct(),
    by = c("RS_CODE", "ReSurvey site", "ReSurvey plot")
  ) %>%
  filter(database == "EVA_RS") %>% # Keep only EVA ReSurvey
  # So far remove cases where ReSurvey plot is NA
  filter(!is.na(`ReSurvey plot`)) %>%
  # Define a group for each ReSurvey plot
  group_by(RS_CODE, `ReSurvey site`, `ReSurvey plot`) %>%
  # For each ReSurvey plot, keep only the last year
  slice_max(order_by = year, n = 1) %>%
  ungroup()  # Ungroup to return a regular tibble
```

Check that there is only one record for each year for each plot: OK.

```{r}
db_resurv_last %>%
  # Define a group for each ReSurvey plot and year
  group_by(RS_CODE, `ReSurvey site`, `ReSurvey plot`, year) %>%
  # Count the number of observations for each group
  summarise(count = n()) %>%
  # Filter groups with more than one observation
  filter(count > 1)
```


```{r, include=FALSE}
# Combine the two databases: EVA and EVA ReSurvey (with only last resurvey)
db_resurv_last_and_EVA <- bind_rows(db_resurv_last, db_EVA_updated_clean)
```

## Availabilty

```{r}
grid.arrange(
  # EVA + EVA ReSurvey (all resurveys)
  ggplot(db_resurv_and_EVA %>% 
           # Create column to show availability of location uncertainty
           mutate(loc_uncertainty_available = 
                  ifelse(is.na(`Location uncertainty (m)`), "N", "Y")),
         aes(loc_uncertainty_available, fill = database)) +
    geom_bar() +
    labs(y = "Number of observations") +
    theme(legend.position = "top") +
    ggtitle("Availability of location uncertainty\nEVA + EVA ReSurvey (all resurveys)"),
  # EVA + EVA ReSurvey (last resurvey)
  ggplot(db_resurv_last_and_EVA %>%
           # Create column to show availability of location uncertainty
           mutate(loc_uncertainty_available =
                    ifelse(is.na(`Location uncertainty (m)`), "N", "Y")),
         aes(loc_uncertainty_available, fill = database)) +
    geom_bar() +
    labs(y = "Number of observations") +
    theme(legend.position = "top") +
    ggtitle("Availability of location uncertainty\nEVA + EVA ReSurvey (last resurvey)"),
  ncol =2)
```

## Distribution when available

```{r fig.height=5, fig.width=11}
grid.arrange(
  ggplot(db_resurv_and_EVA %>%
           # Create column to show two categories for location uncertainty
           mutate(
             Location_uncertainty_range = factor(
               case_when(
                 `Location uncertainty (m)` <= 30 ~ "0 – 30 m",
                 `Location uncertainty (m)` <= 200 ~ "30 m – 200 m",
                 `Location uncertainty (m)` <= 1000 ~ "200 m - 1 km",
                 `Location uncertainty (m)` <= 5000 ~ "1 km - 5 km",
                 `Location uncertainty (m)` <= 10000 ~ "5 km - 10 km",
                 is.na(`Location uncertainty (m)`) ~ "NA",
                 TRUE ~ "10 km +"
                 ), 
               levels = c("0 – 30 m", "30 m – 200 m", "200 m - 1 km", 
                          "1 km - 5 km", "5 km - 10 km", "10 km +", "NA"))
             ),
         aes(Location_uncertainty_range, fill = database)) +
    geom_bar(aes(y = (..count..) / sum(..count..) * 100)) +
    labs(y = "% of observations") +
    theme(legend.position = "top") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    ggtitle("Range of location uncertainty\nEVA + EVA ReSurvey (all resurveys)"),
  ggplot(db_resurv_last_and_EVA %>%
           # Create column to show two categories for location uncertainty
           mutate(
             Location_uncertainty_range = factor(
               case_when(
                 `Location uncertainty (m)` <= 30 ~ "0 – 30 m",
                 `Location uncertainty (m)` <= 200 ~ "30 m – 200 m",
                 `Location uncertainty (m)` <= 1000 ~ "200 m - 1 km",
                 `Location uncertainty (m)` <= 5000 ~ "1 km - 5 km",
                 `Location uncertainty (m)` <= 10000 ~ "5 km - 10 km",
                 is.na(`Location uncertainty (m)`) ~ "NA",
                 TRUE ~ "10 km +"
                 ), 
               levels = c("0 – 30 m", "30 m – 200 m", "200 m - 1 km", 
                          "1 km - 5 km", "5 km - 10 km", "10 km +", "NA"))
             ),
         aes(Location_uncertainty_range, fill = database)) +
    geom_bar(aes(y = (..count..) / sum(..count..) * 100)) +
    labs(y = "% of observations") +
    theme(legend.position = "top") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    ggtitle("Range of location uncertainty\nEVA + EVA ReSurvey (last resurvey)"),
  ncol = 2)
```

## Location method

```{r}
ggplot(db_resurv_and_EVA %>%
         filter(database == "EVA") %>%
         # Create column to show two categories for location uncertainty
         mutate(
           Location_uncertainty_range = factor(
             case_when(
               `Location uncertainty (m)` <= 30 ~ "0 – 30 m",
               `Location uncertainty (m)` <= 200 ~ "30 m – 200 m",
               `Location uncertainty (m)` <= 1000 ~ "200 m - 1 km",
               `Location uncertainty (m)` <= 5000 ~ "1 km - 5 km",
               `Location uncertainty (m)` <= 10000 ~ "5 km - 10 km",
               is.na(`Location uncertainty (m)`) ~ "NA",
               TRUE ~ "10 km +"
               ), 
             levels = c("0 – 30 m", "30 m – 200 m", "200 m - 1 km", 
                        "1 km - 5 km", "5 km - 10 km", "10 km +", "NA")),
           location_method = ifelse(
             `Location method` %in% c("04", "01", "06", "02", "08", "07", "05",
                                      "2", "03", "4"), 
             "Codes", `Location method`)
           ),
       aes(Location_uncertainty_range, fill = location_method)) +
  geom_bar(aes(y = (..count..) / sum(..count..) * 100), 
           position = position_dodge(width = 0.9)) +
  labs(y = "% of observations") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("Range of location uncertainty\nEVA (without ReSurvey)")
ggplot(db_resurv_and_EVA %>%
         filter(database == "EVA_RS") %>%
         # Create column to show two categories for location uncertainty
         mutate(
           Location_uncertainty_range = factor(
             case_when(
               `Location uncertainty (m)` <= 30 ~ "0 – 30 m",
               `Location uncertainty (m)` <= 200 ~ "30 m – 200 m",
               `Location uncertainty (m)` <= 1000 ~ "200 m - 1 km",
               `Location uncertainty (m)` <= 5000 ~ "1 km - 5 km",
               `Location uncertainty (m)` <= 10000 ~ "5 km - 10 km",
               is.na(`Location uncertainty (m)`) ~ "NA",
               TRUE ~ "10 km +"
               ), 
             levels = c("0 – 30 m", "30 m – 200 m", "200 m - 1 km", 
                        "1 km - 5 km", "5 km - 10 km", "10 km +", "NA")),
           location_method = ifelse(
             `Location method` %in% c("04", "01", "06", "02", "08", "07", "05",
                                      "2", "03", "4"), 
             "Codes", `Location method`)
           ),
       aes(Location_uncertainty_range, fill = location_method)) +
  geom_bar(aes(y = (..count..) / sum(..count..) * 100), 
           position = position_dodge(width = 0.9)) +
  labs(y = "% of observations") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("Range of location uncertainty\nEVA ReSurvey (all resurveys)")
ggplot(db_resurv_last_and_EVA %>%
           filter(database == "EVA_RS") %>%
           # Create column to show two categories for location uncertainty
           mutate(
             Location_uncertainty_range = factor(
               case_when(
                 `Location uncertainty (m)` <= 30 ~ "0 – 30 m",
                 `Location uncertainty (m)` <= 200 ~ "30 m – 200 m",
                 `Location uncertainty (m)` <= 1000 ~ "200 m - 1 km",
                 `Location uncertainty (m)` <= 5000 ~ "1 km - 5 km",
                 `Location uncertainty (m)` <= 10000 ~ "5 km - 10 km",
                 is.na(`Location uncertainty (m)`) ~ "NA",
                 TRUE ~ "10 km +"
                 ), 
               levels = c("0 – 30 m", "30 m – 200 m", "200 m - 1 km", 
                          "1 km - 5 km", "5 km - 10 km", "10 km +", "NA")),
             location_method = ifelse(
               `Location method` %in% c("04", "01", "06", "02", "08", "07",
                                        "05", "2", "03", "4"),
               "Codes", `Location method`)
             ),
         aes(Location_uncertainty_range, fill = location_method)) +
    geom_bar(aes(y = (..count..) / sum(..count..) * 100), 
             position = position_dodge(width = 0.9)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(y = "% of observations") +
    ggtitle("Range of location uncertainty\nEVA ReSurvey (last resurveys)")
```

```{r, include=FALSE}
# Session info
sessionInfo()
```