---
title: "Script to compare my update of db_resurv and Stephan's update"
author: "Alicia Valdés"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
---

# Load libraries

```{r}
library(readr)
library(dplyr)
library(here)
library(readxl)
library(stringr)
library(tidyverse)
```

# Read the data

```{r}
# My update
db_resurv_A <- read_tsv(
  here("data", "clean", "db_resurv_updated_clean.csv"))

# Stephan's update
db_resurv_S <- read_tsv(here("..", "..", "..", "Data", "MOTIVATE_data", "Data",
                             "EVA and ReSurvey Data", "Raw data",
                             "200_MOTIVATE20240412_header_notJUICE_with_precise_coordinates.csv"))
```

# Keep same columns in my update and Stephan's update

```{r}
db_resurv_A <- db_resurv_A %>%
  select(-starts_with("EUNIS")) %>%
  select(-c(70:83, 85:93))
```

# Compare the two updates

```{r}
# Join the tibbles by PlotObservationID
joined_tibbles <- db_resurv_A %>%
  inner_join(db_resurv_S, by = "PlotObservationID", suffix = c("_A", "_S"))
```

```{r}
nrow(joined_tibbles %>% filter(country_new != Country_S) %>%
       select(PlotObservationID, country_new, Country_S))
nrow(joined_tibbles %>% filter(Country_A != Country_S) %>%
       select(PlotObservationID, Country_A, Country_S))
```

Differences in Country are not so important!

```{r}
nrow(joined_tibbles %>% filter(`Expert System_A` != `Expert System_S`) %>%
       select(PlotObservationID, `Expert System_A`, `Expert System_S`))
print(
  joined_tibbles %>% filter(`Expert System_A` != `Expert System_S`) %>%
    select(PlotObservationID, `Expert System_A`, `Expert System_S`) %>%
    count(`Expert System_A`, `Expert System_S`),
  n = 100)
```

OK (just because I removed ! etc.).

```{r}
# Define the columns you want to compare
columns_to_compare <- c("ReSurvey plot (Y/N)", "ReSurvey project",
                        "ReSurvey site", "ReSurvey plot",
                        "ReSurvey observation", "RS_CODE", "RS_TIME",
                        "Location uncertainty (m)", "Dataset","Longitude",
                        "Latitude")

# Use map to iterate over the columns and count differences
difference_counts <- map_int(columns_to_compare, ~ {
  nrow(joined_tibbles %>%
         filter(!!sym(paste0(.x, "_A")) != !!sym(paste0(.x, "_S"))) %>%
         select(PlotObservationID, !!sym(paste0(.x, "_A")), !!sym(paste0(.x, "_S"))))
})

# Create a named vector for better readability
names(difference_counts) <- columns_to_compare

# View the counts
print(difference_counts)
```

```{r}
joined_tibbles %>% filter(`ReSurvey site_A` != `ReSurvey site_S`) %>%
       select(PlotObservationID, `ReSurvey site_A`, `ReSurvey site_S`)
```

Remark in Ilona's file: CZ_0001 - Repeated RS_plot should be Slana1.

Change ReSurvey site from "Slana1" to "Slana".

Remark in Ilona's file: SI_0002a - error in RS_SITE changed for RF.

Change ReSurvey site FS to RS.

```{r}
joined_tibbles %>% filter(`ReSurvey plot_A` != `ReSurvey plot_S`) %>%
       select(PlotObservationID, `ReSurvey plot_A`, `ReSurvey plot_S`)
```

Some changes I remember doing, others not.

```{r}
joined_tibbles %>% filter(`ReSurvey observation_A` != `ReSurvey observation_S`) %>%
       select(PlotObservationID, `ReSurvey observation_A`, `ReSurvey observation_S`)
```

Looks like parsing problems maybe?

```{r}
joined_tibbles %>% filter(`Location uncertainty (m)_A` !=
                            `Location uncertainty (m)_S`) %>%
       select(PlotObservationID, RS_CODE_A, `Location uncertainty (m)_A`,
              `Location uncertainty (m)_S`)
```

CZ_0019_048 - Need to update Longitude, Latitude and Location uncertainty (m) (coded as PRECISION).

```{r}
joined_tibbles %>% filter(Longitude_A != Longitude_S) %>%
       select(PlotObservationID, RS_CODE_A, Longitude_A, Longitude_S) %>%
  distinct(RS_CODE_A)
```

```{r}
joined_tibbles %>% filter(Latitude_A != Latitude_S) %>%
       select(PlotObservationID, RS_CODE_A, Latitude_A, Latitude_S) %>%
  distinct(RS_CODE_A)
```

# Conclusion

Using my own version from now on, did not notify Stephan or Ilona about these inconsistencies, check again when Ilona provides new version and report.

# Session info

```{r}
sessionInfo()
```